---
title: "Network dependence test"
author: "Youjin Lee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Generate network dependence

We provide two data generative process for generating network dependent outcomes--network dependence due to (1) direct transmission and (2) latent variable dependence.

Under direct transmission, `peer.process` provides time-evolving outcomes where $Y^{t}_{i}$ depends on $\{ Y^{t}_{j} : j \mbox{ is an adjacent peers of } i \}$. 

```{r, message = FALSE, warning = FALSE}
library(knitr)
library(netdep)
library(MASS)
library(mvrtn)
library(igraph)
library(igraphdata)
# generate network
G = latent.netdep(n.node = 200, rho = 0.2, dep.factor = -1)
A = as.matrix(get.adjacency(G))
outcomes = peer.process(A, max.time = 3, mprob = 0.6, epsilon = 0.1)
names(outcomes)
```


```{r}
result0 = make.permute.moran(A, outcomes$time0, np = 500)
result1 = make.permute.moran(A, outcomes$time1, np = 500)
result2 = make.permute.moran(A, outcomes$time2, np = 500)
result3 = make.permute.moran(A, outcomes$time3, np = 500)
```

```{r}
kable(cbind( c("t=0", "t=1", "t=2", "t=3"),
  round(c(result0$moran, result1$moran, result2$moran, result3$moran),2),
  round(c(result0$pval.permute, result1$pval.permute, result2$pval.permute, result3$pval.permute),2)), row.names = NA, col.names = c("Transmission time", "Moran's I", "P-value (permutation)"), 
      caption = "Direct transmission")
```

Note from the above table that as `t` increases, Moran's $I$ statistic increases and p-value decreases. 

Next we generate network dependent continuous outcomes via (2) latent variable dependence, and then transform those into categorical outcomes, where $\Phi$ statistic can be applicable to test network dependence.



## Real Data

```{r}
library(igraphdata)
data(karate)
A = as.matrix(as_adjacency_matrix(karate, attr= "weight", sparse = T)) # weighted adjacency matrix
Y = V(karate)$Faction
result = make.permute.moran(A, Y, np = 500)
print(result$moran)
print(result$pval.permute)
```


```{r}
data(UKfaculty)
A = as.matrix(as_adjacency_matrix(UKfaculty, attr= "weight", sparse = T)) # weighted adjacency matrix
Y = V(UKfaculty)$Group
table(Y)
result = make.permute.Phi(A, Y, np = 500)
print(result$phi)
print(result$pval.permute)
```


More details on the method and applications can be found in [Arxiv paper](https://arxiv.org/pdf/1710.03296.pdf).
